---
title: "The left-right positions of British MPs inferred from a survey of local councillors"
published: "Working paper"
code-repo: "Access the code, data, and analysis at <>"
correspondence-prefix: "Correspondence concerning this article should be addressed to"
author:
  - name: Chris Hanretty
    email: chris.hanretty@rhul.ac.uk
    orcid: 0000-0002-3948-3914
    title: Professor of Politics
    affiliations: 
      - id: rhul
        name: "Royal Holloway, University of London"
        department: "Department of Politics, International Relations and Philosophy"
        address: McCrea 1-08
        city: Egham
        region: Surrey
        country: United Kingdom
        postal-code: TW20 0EX
    attributes:
      corresponding: true
  - name: Vasil Lazarov
    email: vasil.lazarov@survation.com
    title: PhD student
    affiliations:
      - id: survation
        name: Survation
      - ref: rhul
abstract: "We present local councillors in England, Wales and Scotland with pairwise comparisons between MPs in their local area and two anchor MPs and ask them to identify the more left-wing MP on economic issues (first and second survey waves) and on cultural issues (second wave only). We infer MPs' positions on the economic and cultural left-right dimension from the responses to these pairwise choices. The estimates of MPs' positions have good face validity, but our estimates or MPs' positions on the cultural dimension are nearly identical to our estimates of the MPs' positions on the economic dimension, suggesting that knowledgeable respondents cannot distinguish MPs' positions on these two dimensions. Our estimates can be used to study the effects of MPs' position on behaviour in the party and in the legislature. "
thanks: "Thanks to Survation for hosting these questions on their omnibus survey of local councillors. Particular thanks to David Izamoje for setting up the script seen by councillors. "
additional-info: |
  We have no known conflict of interest to disclose.
keywords:
  - expert survey
  - ideal points
  - ordinal regression
date: January 2, 2025
bibliography: biblio.bib
mainfont: "TeX Gyre Pagella"
sansfont: "TeX Gyre Heros"
mathfont: "Asana Math"
prefer-html: true
execute:
  echo: false
  message: false
  warning: false
format:
  hikmah-pdf:
    number-sections: true
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{amsmath}
##  hikmah-manuscript-docx: default
---

<!--

Structure needs to be:

 - introduction
 - wave 1:
 - wave 2: two dimensions
 - combining multiple waves
 
-->


```{r}
here::i_am("article/article.qmd")
suppressPackageStartupMessages(library(tidyverse))
library(knitr)
library(ggbeeswarm)
library(kableExtra)
library(stringr)
library(scales)
```

# Introduction

The measurement of actors' positions in political space is a key task
of political science. Most effort has been spent on estimating the
positions of parties in a one-dimensional, left-right, political
space; but for polities which use candidate-centred electoral systems,
or which have politically relevant within-party disagreement, knowing
the positions of individual legislators is also
important. Unfortunately, measuring individual legislators' positions
is difficult outside of a small number of countries. Most legislators
do not issue personal manifestos or policy platforms
[cf. @catalinac2018positioning], meaning we cannot scale legislators
in the same way as (for example) the Comparative Manifesto Project
scales political parties [@volkens2013mapping]. Although legislators
do talk and vote in parliaments, these acts are often strategic:
scaling techniques used successful to analyse congressional roll-calls
can fail to recover left-right positions where there are strong
government-opposition dynamics and where extremes of left and right
join together to vote against the middle [@spirling2007uk].[^edms] The
analysis of legislative speech generally also tends to recover
government-opposition dynamics more than left-right position
[@lauderdale2016measuring].[^topics] Whilst expert surveys have proved
tremendously useful in the study of party systems, few political
scientists would be able to place hundreds of MPs on a 0-10 scale with
as much equanimity as they place parties. Where political scientists
do make expert assessments of individual MPs, these assessments are
typically coarse-grained [@heppell2013cameron].

[^edms]: Some non-strategic legislative behaviours can be analysed
    [@kellermann2012estimating], but these are becoming less popular
    (and thus less informative)

[^topics]: Latent positions on specific issues can be recovered, but
    this requires a careful selection both of texts and reference
    points. See @o2022transformation. 

In this note, we provide estimates of the economic left-right
positions of British MPs in the 2019-2024 and 2024-
parliaments. Estimating the positions of British legislators is
important because within-party disagreement in Britain has been
increasing over time, and because MPs' preferences have strongly
conditioned or determined the identity of four of the last five
British prime ministers.[^pms] We estimate positions on the basis of a
survey of local councillors. We presented councillors with up to six
pairwise comparisons between MPs in their local area and two "anchor"
MPs (Prime Minister Rishi Sunak and Labour leader Sir Keir
Starmer). This pairwise approach has previously been used by
@breunig2021measuring and @hopkins2022trump; our application uses a
richer set of response categories and presented only local
comparisons. 

[^pms]: Theresa May was elected Conservative party leader without a
    membership vote after four ballots of Conservative MPs; Boris
    Johnson became leader by beating Jeremy Hunt in a membership vote
    after eight other candidates were eliminated by MPs; Liz Truss
    became leader by beating Rishi Sunak in a membership vote after
    six other candidates were eliminated by MPs; Rishi Sunak
    subsequently became leader without any membership vote or ballot
    amongst MPs. Sir Keir Starmer became Labour leader after a
    membership ballot, but was prohibitive favourite after receiving
    twice as many nominations from parliamentary colleagues as the
    next best-placed candidate (Rebecca Long-Bailey).
	
Although the details of our survey differed slightly between the two
survey waves (conducted in late 2023 and 2024 respectively), we asked
councillors to pick the more right-wing on economic and cultural
issues, choosing between five different response options. We analyze
these ordinal responses using a Bayesian ordinal logistic regression
[@burkner2019ordinal] with symmetric thresholds [@johnson2003use]. The
use of Bayesian methods allows us to work directly with the
probability that a named MP is more left- or right-wing than a
comparison MP and calculate measures of uncertainty for derived
statistics such as rank-order. Our estimates should prove useful to
researchers interested in parliamentary representation in the United
Kingdom and more generally.

There are five sections to this note. In section two, we provide a
description of the survey data, giving details on the total number of
respondents, the total number of pairwise comparisons, and how we
presented respondents with MPs. In section three, we describe the
model that we use. Because we have to deal with positions which are
structured by party, time and different dimensions, this section is
necessarily long and detailed. In section four, we discuss the
estimates of MPs' positions, with a view to establishing face validity
and to describing how confident the model can be about the positions
of particular MPs. In the final section we discuss some of the
limitations of the research.

# Description of the survey

This note is based on two surveys of local councillors in England,
Wales and Scotland.[^ni] The first survey was conducted between the
7th August and 3rd September 2023, and asked about members of the
2019 - 2024 parliament. The second survey was conducted between 14th
and 24th October 2024, and asked about members of the parliament
elected in the 2024 general election. In total, 1486 councillors
responded to the first survey, and 1027 to the second survey. 

[^ni]: We also surveyed local councillors in Northern Ireland, but
    including Northern Irish MPs in the analysis created problems of
    model fit, possibly because Northern Irish respondents are
    evaluating the anchor politicians (Sir Keir Starmer and Rishi
    Sunak) in a different way to respondents in the other three
    countries. 

Rather than asking respondents to evaluate a single politician by
placing them on a scale, we asked respondents to compare politicians
two at a time. In the first survey, respondents were presented with
two MPs and asked to pick which MP was the *more left-wing* on
*economic* issues. Respondents had six different response options:
they could say that the first-named MP was much more left-wing; that
the first-named MP was somewhat more left-wing; that the two MPs were
about the same; that the second-named MP was somewhat more left-wing;
that the second-named MP was much more left-wing, or that they did not
know enough to say. Although we have described these response options
in terms of the first- and second-named MP, the response options seen
by respondents featured the actual names of the MPs. Respondents to
the first wave were asked to make up to six comparisons. The number of
comparisons could be fewer than six if there were few MPs in the
respondent's local area, as defined below. 

In the second survey, we changed the question wording. We did this
because we wanted to ask about MPs' positions on the cultural
dimension as well as the economic dimension. We began the survey by
asking respondents to think about economic issues. We then presented
respondents with up two six comparisons and asked them to say whether
the first-named MP was much more *economically conservative*, somewhat
more economically conservative, and so on. We then asked respondents
to consider cultural issues. Once again we presented respondents with
MPs to compare, but asked them to say which MPs was more *culturally
conservative*. We asked respondents to pick the more *conservative* of
the two MPs because we expected that some respondents would interpret
"left" and "right" to have economic content only (such that we could
not ask them to pick the more left-wing MP on cultural issues), and
because the word "conservative" picks out a right-leaning position on
both the economic and social dimensions, which is not true of words
like *liberal*, which (in British usage) can imply a laissez-faire
economic policy. 

We have described respondents as comparing MPs "in their local
area". Specifically, we match each respondent's local council to the
"upper tier local authority area" (UTLA). The meaning of UTLA differs
between country. For English respondents in district councils, this is
their county council. For English respondents in unitary authorities,
this is the unitary authority itself. For respondents in Scotland and
Wales, it is simply the local authority. For each UTLA we identify all
Westminster constituencies which entirely or partly covered by that
area. In the second wave of the survey, where a UTLA was represented
by a single Westminster constituency, we grouped that UTLA with the
contiguous UTLA represented by the fewest Westminster MPs. We then
generate all pairwise comparisons of MPs serving those constituencies,
*plus* the two anchor MPs, Sir Keir Starmer and Rishi Sunak. We ask
include these two anchor MPs in order to place different local
comparisons on a national scale: all we would have would be a series
of regional orderings of MPs with no overlap. We then keep all
comparisons between local MPs, no comparisons between Sunak and
Starmer, and two comparisons involving Sunak and Starmer.

A worked example may help. Consider a respondent councillor in Castle
Point. The UTLA for Castle Point District Council is Essex County
Council. Essex County council includes sixteen Westminster
constituencies.[^southend] With two anchor MPs added, there are 153
pairwise combinations of MPs ignoring the order of presentation. Of
these, 120 are comparisons between Essex MPs; one is a comparison
between Sunak and Starmer, and thirty-two are comparisons involving
Sunak and Starmer. We keep all 120 comparisons between Essex MPs and
two comparisons involving Sunak and Starmer. Of these 122 comparisons,
we randomly show respondents six comparisons. 

[^southend]: Southend-on-Sea and Thurrock, as unitary authorities, are
    their own UTLAs.

```{r}
svy <- dat <- readRDS(file = here::here("working",
                                 "combined_data.rds"))
n_w1 <- sum(dat$wave == 1)
n_w2 <- sum(dat$wave == 2)
```

Respondents to the first wave of the survey expressed `r n_w1`
pairwise comparisons between MPs, excluding don't know
responses. Respondents to the second wave expressed `r n_w2` pairwise
comparisons excluding don't know responses. The number of pairwise
comparisons in the second wave is larger because respondents were
presented with six comparisons on each dimension, and because in the
second wave we grouped some UTLAs together. 

```{r}
#| label: tbl-dvw1
#| tbl-cap: Response distribution in the first wave, asking about economic issues. Order of responses is reversed from how it appeared in the original survey.

tab_1_df <- dat |>
    filter(wave == 1) |>
    count(dv) |>
    mutate(Response = dplyr::recode(dv,
                                    "5" = "[First named MP] is much more left-wing",
                                    "4" = "[First named MP] is somewhat more left-wing",
                                    "3" = "[First named MP] and [second named MP] are about the same",
                                    "2" = "[Second named MP] is somewhat more left-wing",
                                    "1" = "[Second named MP] is much more left-wing")) |>
    dplyr::select(Response, Y = dv, n)

kable(tab_1_df)
```


```{r}
#| label: tbl-econdvw2
#| tbl-cap: Response distribution in the second wave, asking about economic issues

tab_2_df <- dat |>
    filter(wave == 2) |>
    filter(dim == "econ") |>
    count(dv) |>
    mutate(Response = dplyr::recode(dv,
                                    "1" = "[First named MP] is much more economically conservative",
                                    "2" = "[First named MP] is somewhat more economically conservative",
                                    "3" = "[First named MP] and [second named MP] are about the same",
                                    "4" = "[Second named MP] is somewhat more economically conservative",
                                    "5" = "[Second named MP] is much more economically conservative")) |>
    dplyr::select(Response, Y = dv, n)

kable(tab_2_df)
```

```{r}
#| label: tbl-cultdvw2
#| tbl-cap: Response distribution in the second wave, asking about cultural issues

tab_3_df <- dat |>
    filter(wave == 2) |>
    filter(dim == "cult") |>
    count(dv) |>
    mutate(Response = dplyr::recode(dv,
                                    "1" = "[First named MP] is much more culturally conservative",
                                    "2" = "[First named MP] is somewhat more culturally conservative",
                                    "3" = "[First named MP] and [second named MP] are about the same",
                                    "4" = "[Second named MP] is somewhat more culturally conservative",
                                    "5" = "[Second named MP] is much more culturally conservative")) |>
    dplyr::select(Response, Y = dv, n)

kable(tab_3_df)
```


The distribution of responses is shown in @tbl-dvw1, @tbl-econdvw2 and
@tbl-cultdvw2. The order of responses in @tbl-dvw1 has been reversed
because wave 1 asked respondents to identify the more left-wing MP and
wave 2 asked respondents to identify the more (economically or
socially) conservative. The modal response is that the MPs being
compared are roughly the same, and extreme responses ("much more") are
always less commmon than less extreme responses ("somewhat
more"). There is some evidence in the second wave for an order effect,
such that MPs who are listed second in the comparison are more likely
to be picked than the first-listed MP in the comparison.

<!-- response rate; don't knows -->

<!-- distribution of responses -->

<!-- distribution of responses across MPs -->

```{r}
uniq_mps <- unique(c(dat$wikidata.A,
                      dat$wikidata.B))
n_mps <- length(uniq_mps)

ni_seats <- c("Belfast East", "Belfast North",
              "Belfast South and Mid Down",
              "Belfast West",
              "East Antrim",
              "East Londonderry",
              "Fermanagh and South Tyrone",
              "Foyle",
              "Lagan Valley",
              "Mid Ulster",
              "Newry and Armagh",
              "North Antrim",
              "North Down",
              "South Antrim",
              "South Down",
              "Strangford",
              "Upper Bann",
              "West Tyrone")

all_w1 <- read.csv(here::here("data",
                              "all_wave1_mps_per_wikidata.csv"))

all_w1 <- all_w1 |>
    filter(!is.element(constituencyLabel, ni_seats))

w1_missing <- setdiff(all_w1$item, uniq_mps)

n_w1_misses <- length(w1_missing)

all_w2 <- read.csv(here::here("data",
                              "all_wave2_mps_per_wikidata.csv"))

all_w2 <- all_w2 |>
    filter(!is.element(constituencyLabel, ni_seats))

w2_missing <- setdiff(all_w2$item, uniq_mps)

n_w2_misses <- length(w2_missing)

n_keir <- sum(c(dat$wikidata.A,
                dat$wikidata.B) == "http://www.wikidata.org/entity/Q6383803")

n_rishi <- sum(c(dat$wikidata.A,
                 dat$wikidata.B) == "http://www.wikidata.org/entity/Q20055561")

med_appear <- median(table(c(dat$wikidata.A, dat$wikidata.B)))

```

In total respondents expressed opinions about 953 MPs. There were 
`r n_w1_misses` British MPs who were in office at the time of the first
wave who are not featured in this data. There are `r n_w2_misses` MPs
who were in office at the time of the second wave who are not featured
in this data. MPs may not feature in the data because they were not
asked about or because no respondents gave usable responses. MPs who
are not featured in the data do not have any estimated position.

Considering just the MPs who are featured in the data, some MPs
feature much more than others. This is particularly true for the two
anchor MPs, Sir Keir Starmer and Rishi Sunak, who appear `r n_keir`
and `r n_rishi` times respectively. The average (median) MP features
in the data `r med_appear` times.


# Description of the model

In this section we describe the model that we use to connect
councillors' responses to MP positions. As a reminder: councillor's
responses are an ordinal variable with values 1 to 5, where a value of
5 corresponds to a judgement that the second-named MP is much more
(economically or culturally) conservative than the first-named MP, and
a value of 1 corresponds to a judgement that the *first-named* MP is
much more (economically or culturally) conservative than the
second-named MP. Responses from the first-wave, which asked about the
more left-wing MP, have been flipped so that they match responses from
the second wave. We "explain" these ordinal responses through a latent
continuous variable, which is turned into an ordinal variable through
a series of cutpoints. We use $\mu_i$ to refer to the latent
continuous variable, where the subscript $i$ runs from one to 13680
(the total number of judgements made by councillors). We use
$\symbf{\tau}$ to refer to the cutpoints. Because there are five
response categories, there are four (five minus one) cutpoints between
these categories.

We connect the ordinal response to the latent continuous variable by
using a cumulative logit link model. A cumulative link model models
the probability that the response will fall in category $k$ or lower:

$$
Pr(Y_i \le k) = F(\tau_{k - 1} - \mu_i)
$$

where $F()$ is some function which takes in the difference between the
cutpoint and the latent parameter (which can be any real number,
positive or negative), and returns a probability between zero and
one. We use the cumulative distribution function for the logistic
distribution, which makes our model a cumulative logit link model. The
use of a logit link rather than (say) a probit link does not affect
our estimates except for changing the scale of the estimates. 

In a cumulative link model, the cutpoints $\symbf{\tau}$ can either be
flexible or structured. We use structured cutpoints. Specifically, we
ensure that the cutpoints are symmetric, such that the distance
between cutpoints one and two is the same as the distance between
cutpoints three and four. We use symmetric thresholds because of the
symmetry in our question format: whether an MP appears as the
first-named or second-named MP is random, and so the distance between
"much more" and "somewhat more" should not be different depending on
whether we face the actual comparison, or the same comparison but with
the order reversed. Formally, we set

$$
\begin{aligned}
\symbf{\tau} = [ & \tau^*_1,  \\
& \tau^*_1 + \delta_1, \\
& \tau^*_1 + \delta_1 + \delta_2, \\
& \tau^*_1 + 2 \cdot \delta_1 + \delta_2 
]
\end{aligned}
$$

where $\delta_1$ and $\delta_2$ are different increments or
"spacers". Using spacers in this way is not the only way of ensuring
symmetry [@christensen2019clm], but we find it is the easiest way. A
visual representation of this setup is given in @fig-thresh, which
shows the thresholds and the increments used.


```{r, fig = TRUE, dpi = 300}
#| fig-width: 9
#| fig-height: 4
#| label: fig-thresh
#| fig-cap: "Illustration of the symmetric thresholds used"
thresh0 <- -1
delta1 <- 1/2
delta2 <- 1
thresh <- c(thresh0,
            thresh0 + delta1,
            thresh0 + delta1 + delta2,
            thresh0 + delta1 + delta2 + delta1)

### Create the graph area
par(mar = c(4, 2, 2, 2))
plot(x = c(-3, 3),
     y = c(0, 0.65),
     type = "n",
     xlab = expression("Linear predictor"~mu),
     ylab = "",
     axes = FALSE)
axis(1)

### Add the curve
curve(dnorm, from = -3, to = 3, lwd = 2, add = TRUE)
### Add the thresholds: lines from 0 to 0.4
ytop <- 0.4
for (i in 1:4) {
    lines(x = c(thresh[i], thresh[i]),
          y = c(0, 0.4))
}
text(x = thresh[1], y = 0.4, expression(tau[1]), pos = 3)
text(x = thresh[2], y = 0.4, expression(tau[2]), pos = 3)
text(x = thresh[3], y = 0.4, expression(tau[3]), pos = 3)
text(x = thresh[4], y = 0.4, expression(tau[4]), pos = 3)

### Add on the DV descriptors at the top
ytop <- 0.625

text(x = -1.5, y = ytop, "Y = 1")
### Interiors
text(x = mean(thresh[1:2]), y = ytop, "Y = 2")
text(x = mean(thresh[2:3]), y = ytop, "Y = 3")
text(x = mean(thresh[3:4]), y = ytop, "Y = 4")
###
text(x = 1.5, y = ytop, "Y = 5")

### Add on the DV text
ytop <- 0.6

text(x = -1.5, y = ytop, "[First named]\nmuch more\nl-w", cex = 0.8, adj = c(0.5, 1))
### Interiors
text(x = mean(thresh[1:2]), y = ytop,
     "...somewhat\nmore l-w...",
     cex = 0.8, adj = c(0.5, 1))
text(x = mean(thresh[2:3]), y = ytop,
     "...about the\nsame...",
     cex = 0.8, adj = c(0.5, 1))
text(x = mean(thresh[3:4]),
     y = ytop, "[First-named]\nsomewhat more\nr-w",
     cex = 0.8, adj = c(0.5, 1))
###
text(x = 1.5, y = ytop, "... much\nmore r-w", cex = 0.8, adj = c(0.5, 1))

### Add on the symmetric distances
ymid <- 0.15
arr_length <- unit(0.15, "inches")
arrows(x0 = thresh[1], y0 = ymid, x1 = thresh[2], y1 = ymid, code = 3, length = arr_length)
arrows(x0 = thresh[2], y0 = ymid + 0.05, x1 = thresh[3], y1 = ymid + 0.05, code = 3, length = arr_length)
arrows(x0 = thresh[3], y0 = ymid, x1 = thresh[4], y1 = ymid, code = 3, length = arr_length)

### Add the text above
text(x = mean(thresh[1:2]), y = ymid + 0.025, expression(delta[1]))
text(x = mean(thresh[2:3]), y = ymid + 0.075, expression(delta[2]))
text(x = mean(thresh[3:4]), y = ymid + 0.025, expression(delta[1]))

```


The thresholds explain how we can connect the ordered responses given
by councillors to something continuous, but they do not explain where
that continuous variable $\mu$ comes from. We set up $\mu$ so that it
is equal to the difference between the latent positions of the two MPs
being compared. We use $\theta$ to refer to the latent position of
each MP. Referring to $\theta$ is complicated because
the positions of MPs can be positions of (i) different MPs; (ii) at
different times; (iii) on different dimensions. Additionally,
referring to $\theta$ is complicated because we sometimes wish to the
position of an MP by their index -- for example, the position of MP
number fifty on our list of MPs -- and sometimes by their role in a
comparison -- for example, as the MP listed as the first-named MP on
question forty. Bearing this in mind, we use subscripts and
superscripts so that $\theta^{Econ.}_{j, t}$ refers to the position of
MP $j$ on the economic dimension at time $t$. The subscript $j$ runs
from 1 to 953 (the number of MPs in our combined data); the subscript
$t$ is either 1 (referring to the survey conducted in autum of 2023)
or 2 (referring to the survey conducted in the autumn of 2024). When
we write $\theta^{Cult.}_{A[i], t}$, we refer to the position on the
cultural dimension of the first-named MP in question $i$; when we
write $\theta^{Econ.}_{B[i], t}$ we refer to the position on the
economic dimension of the second-named MP in question $i$.

With this notation, we can then say that

\doublespacing

$$
\mu_i = \left\{ 
  \begin{array}{ c l }
    \theta^{Econ.}_{B[i], t} - \theta^{Econ.}_{A[i], t} & \quad \textrm{if question was about economics} \\
    \theta^{Cult.}_{B[i], t} - \theta^{Cult.}_{A[i], t} & \quad \textrm{otherwise}
  \end{array}
\right.
$$

\singlespacing

In one sense, this completes the connection between the responses
given by councillors and the positions of the MPs on the two
dimensions: we have specified a model which ends in positions of
different MP on two dimensions at different times. However, this is
not the end of the model specification. Just as with the cutpoints, we
chose to introduce certain elements of structure so that the model
makes sense. We chose to capture three types of structure: temporal,
issue, and party. 

Consider time first. Even considering just the economic dimension,
we've posited two values of $\theta$ for each MP -- one value in the
first wave, one value in the second wave. It's very likely that these
two values are close together -- that the distance between the same MP
a year apart is much less than the distance between two MPs chosen at
random but "measured" in the same year. We can model this by making
the MP's position at time two equal to their position at time one,
plus a small shift -- or equivalently, we can take the most recent
position as the baseline, and make positions at previous points in
time equal to their most recent position, minus a small
shift. Formally,

$$
\theta^{Econ.}_{j, 1} = \theta^{Econ.}_{j, 2} + \nu_j
$$

and

$$
\theta^{Cult.}_{j, 1} = \theta^{Cult.}_{j, 2} + \nu_j
$$

where these small shifts are in turn modelled as coming from a common
distribution with a specified standard deviation. 

$$
\nu \sim N(0, 0.1)
$$

By imposing the assumption that the small shifts come from a common
distribution we are able to move back and forth between how large
shifts are on average and how large a shift might be in any particular
case. 

Now consider party. Just as an MP's position in 2024 is likely to
resemble their position in 2023, so too will the position of a Labour
MP resemble that of another Labour MP much more than it does a
Conservative MP. Although we use the Labour and Conservative parties as
examples here, this applies to any party (or group of MPs who can be
treated as though they were a party). We therefore model each MPs'
overall position as a function of a party-specific component, and a
component which is specific to them. That is,

\doublespacing
$$
\theta^{Econ.}_j = \alpha^{Econ.}_j + \gamma^{Econ.}_{P[j]}
$$

$$
\theta^{Cult.}_j = \alpha^{Cult.}_j + \gamma^{Cult.}_{P[j]}
$$

\singlespacing

where $P[j]$ gives the party of legislator $j$, and $\gamma$ refers to
a set of party-specific fixed effects relative to the reference party,
the Conservatives.

Finally, consider associations between the different dimensions. It is
logically possible that each MPs' position on the economic dimension
tells us nothing about their position about their position on the
cultural dimension -- that a free-marketeer is as likely to be a
social libertarian as they are to be a moralizer. At the same time,
there is lots of evidence to suggest that Western European elites (a
category which necessarily includes members of national legislatures)
are very good at subsuming economic and cultural issues under a single
rubric which has to do with equality. Positions on these two
dimensions may therefore be correlated to a greater or lesser
degree. We can incorporate this into the model by having the
idiosyncratic legislator effects be drawn from a multivariate normal
distribution with an estimated covariance matrix. Formally,

$$
\begin{bmatrix}
\alpha^{Econ.} \\
\alpha^{Cult.}
\end{bmatrix} \sim
N\left(
\begin{bmatrix}
0 \\
0 
\end{bmatrix},
\begin{bmatrix}
\sigma_{Econ.} & \sigma_r \\
\sigma_r & \sigma_{Cult.}
\end{bmatrix}
\right)
$$

The parameters in the model are therefore of three types. There are
parameters of primary interest: the positions of MPs on each dimension
at each time. These are in turn built on parameters of secondary
interest, like the party-specific effects and the legislator-specific
effects on each dimension. Finally, there are ancillary parameters,
like the increments over time, and the variances and covariances. The
model is therefore complicated, but it is only as complicated as it
needs to be to model data asking about two dimensions at two different
points in time.

It would in principle be possible to incorporate other elements into
the model in the same way that party has been incorporated into the
model. Constituency-specific features like the share of the
Conservative vote might be included on the basis that Conservative MPs
from an area where the Conservatives are generally popular are more
likely to be right-wing than Conservative MPs who just eked out a
win. However, including these additional covariates raises further
questions (should we expect the Conservative share of the vote to work
in the same way across different elections?) and complicates attempts
to relate MP positions to characteristics of their area (how can we
assess the strength between voting behaviour in the area and MP
positions when voting behaviour in the area is used to construct the
measure of MP positions?). 

## Identifiability and scale

Ordinal logit models can suffer from problems of identifiability when
the continuous predictor $\mu$ is modelled as the difference of two
latent variables, in the sense that different sets of parameter values
can give the same fit to the data. These identifiability issues can be
broken down into issues of scale and location. Problems of scale arise
because we can multiply the latent variables by a constant and divide
the cut-off parameter by that constant, leaving the fit to the data
unchanged. Problems of location arise because we can add on a constant
to all latent variables, leaving the fit to the data unchanged. This
is because the linear predictor depends on the difference between the
latent variables rather than their absolute values. 

The issue of scale is dealt with by placing a prior on the first
threshold centred on -2, so that it would no longer be possible to
multiply the latent parameters by some constant and offset that change
by altering the cut-off parameters. The issue of location is dealt
with by the use of the prior on the legislator-specific component of
legislator's ideal points. This prior is centred on zero, and so
solutions would involve adding large positive or negative constants to
the ideal points are penalized.

Although the scale is set by the model (in the sense that the first
cut-off is drawn from a distribution centred on -2 and priors are
placed on the distribution of MP ideal points), this scale is not
intuitive. In this note, we therefore scale MP positions by
subtracting the minimum, dividing by the maximum, and multiplying
by 100. This places the positions on a 0-100 scale. Because the
minimum and maximum are properties of the sample we collected,
estimates which are transformed in this way will not be comparable
across different and separately modelled surveys. Researchers should
not therefore interpret these positions as though they were absolute
measurements, or imply that a score of 25, or 50, or 75, means that an
MP is "left-wing", "centrist", or "right-wing".


## Prior specification

We place prior distributions on some of these parameters and set some
parameters as fixed. We do this to solve the problems of scale and
location noted above. Specifically, we adopt the following priors for
the cutpoints

$$
\tau_1 \sim N(-2, 0.5)
$$

$$
\delta_1 \sim N^+(0, \sqrt{3})
$$

$$
\delta_2 \sim N^+(0, \sqrt{3})
$$

where $N^+$ indicates a truncated, or half-normal distribution which
only generates positive values. 

We draw party effects from the following common distribution:

$$
\gamma \sim N(0, 4.0)
$$

The most complex priors are the priors placed on the covariance
between legislator specific ideal points. We decompose this covariance
matrix into the product of a scale vector S and a correlation matrix
R, where the prior on the correlation matrix R involves the
Lewandowski-Kurowicka-Joe distribution [@lewandowski2009generating].

$$
\begin{bmatrix} \sigma_{Econ.} & \sigma_r \\ \sigma_r &
\sigma_{Cult.}  \end{bmatrix} \equiv \operatorname{diag}(S) \cdot R \cdot \operatorname{diag}(S) 
$$

$$
S \equiv
\begin{bmatrix}
\varsigma_{Econ.}, \varsigma_{Cult.}
\end{bmatrix}
$$

$$
\varsigma \sim N^{+}(0, 2);
$$

$$
R \sim LJKCorr(4)
$$

A full justification of these priors is given in the appendix.


## Estimation

The model was estimated in `rstan`, the R interface to the Stan
programming language [@rstan; ]. The model was set to run for 2,000 iterations,
of which 1,000 iterations were discarded as warm-up
iterations. Convergence across the four chains, as measured by the
R-hat statistic, was acceptable. 

# Results

In this section we describe the results of the analysis, focusing exclusively on the positions of MPs, rather than any of the secondary and auxiliary parameters in the model. We make seven broad points regarding these positions.

*First,* the left-right ordering of parties (according to the average
position of MPs in that party on the economic dimension) makes
sense. @fig-bees shows a beeswarm plot of MP's left-right position on
the economic dimension for MPs elected in the 2024 - parliament. Each
plotted point corresponds to the posterior mean. Values have been
rescaled so that the maximum value across all MPs and iterations is
100, and so that the minimum value is zero. Parties are ordered by
their average (mean) position on this dimension. MPs from Reform UK
are thus the most right-wing party, and the Greens the most left-wing
party. This matches the rank ordering of parties in the 2024 British
Election Study Expert Survey, *except that* the expert survey places
Plaid Cymru to the left of the Scottish National Party, whilst these
estimates place the SNP to the left. 

@fig-bees is valuable not just because it shows the relative order of
parties, but also relative dispersion of the MPs in that party. When
judged by the *range* of MPs positions, the Labour party is the most
variable; when judged by the *standard deviation* of positions, it is
the Conservative party which is most variable. Another way of saying
the same thing is to note that whilst the Labour party has some
outlying MPs to the extreme left and to the centre, many of its MPs
are clustered around the average within the party.



```{r}
#| fig-width: 9
#| fig-height: 6
#| label: fig-bees
#| fig-cap: "Bee-swarm plot of MPs' left-right positions on the economic dimension for MPs elected in 2024. "

dat <- read.csv(here::here("outputs",
                           "summary_statistics_file.csv"))

plot_df <- dat |>
    filter(!is.na(Party.24)) |>
    filter(Party.24 != "Spk") |>
    filter(Party.24 != "Ind") |>
    mutate(Party = factor(Party.24, ordered = TRUE),
           Party = reorder(Party, econ_mean)) 

### Get count within party to use for alpha
plot_df <- plot_df |>
    group_by(Party) |>
    mutate(n_members = n())

ggplot(plot_df, aes(x = Party, y = econ_mean)) +
    geom_beeswarm(aes(fill = Party, colour = Party, shape = Party,
                      alpha = 1 / n_members),
                  size = 3,
                  cex = 1) +
    scale_x_discrete("Party", labels = function(x) str_wrap(x, width = 10)) +
    scale_y_continuous("Estimated left-right position on the economic dimension") +
    scale_shape_manual("Party",
                      values = c("Con" = 21,
                                 "Lab" = 22,
                                 "LD" = 23,
                                 "SNP" = 24,
                                 "PC" = 25,
                                 "Green" = 21,
                                 "RUK" = 22)) +
    scale_alpha_continuous(guide = "none",
                           range = c(1/3, 1)) + 
    scale_fill_manual("Party",
                  values = c("Con" = "#0087DC",
                             "Lab" = "#E4003B",
                             "LD" = "#FAA61A",
                             "SNP" = "#FDF38E",
                             "PC" = "#005B54",
                             "Green" = "#528D6B",
                             "RUK" = "#12B6CF")) +
    scale_colour_manual("Party",
                      values = c("Con" = "darkgrey",
                                 "Lab" = "darkgrey",
                                 "LD" = "black",
                                 "SNP" = "black",
                                 "PC" = "darkgrey",
                                 "Green" = "darkgrey",
                                 "RUK" = "black"
                                 )) +
    coord_flip() + 
    theme_bw(base_size = 12) +
    theme(legend.position = "bottom")

```


*Second,* individuals at the extremes of the recovered distribution
are all individuals who we would, on the basis of other evidence,
describe as being at the extreme ends of their party.[^extremes]
Whilst @fig-bees is helpful in understand averages, we can gain a
better understanding of these positions by considering individuals MPs
are the left, middle, and right of the distribution. @tbl-extremes
shows eighteen MPs divided equally between these three categories. For
each MP we show their average position (column labelled "Mean" under
"Position"), the 5th and 95th percentiles of the distribution (columns
marked "Lo" and "Hi" under "Position"). We also show equivalent
figures for the *rank* of the MP, where a rank of 1 means that the MP
is the most left-wing and a rank of 614 means that the MP is the most
right-wing of the MPs to feature in the data.

[^extremes]: Once again, it is important to note that "extreme" as
    used here just means "extreme relative to the other MPs asked
    about in the survey", and does not imply extremism in an absolute
    sense.

```{r}
#| label: tbl-extremes
#| tbl-cap: MPs on the left, middle, and right of the recovered dimension. Columns Lo and Hi refer to the 2.5th and 97.5th percentiles of the posterior distribution. 
plot_df <- read_csv(here::here("outputs",
                                 "mpsleftright_full.csv.gz"))

plot_df <- plot_df |>
    filter(dimension == 1) |>
    filter(!is.na(PCON24NM)) |>
    group_by(iter) |>
    arrange(value) |>
    mutate(econ_rank = 1:n()) |>
    ungroup() |>
    group_by(wikidata, DisplayName, Party.24, PCON24NM) |>
    summarize(econ_mean = mean(value.sc),
              econ_lo = quantile(value.sc, 0.05),
              econ_hi = quantile(value.sc, 1 - 0.05),
              actual_rank = round(mean(econ_rank)),
              lo_rank = quantile(econ_rank, 0.05),
              hi_rank = quantile(econ_rank, 1 - 0.05),
              .groups = "drop")

tab_top <- plot_df |>
    arrange(econ_mean) |>
    head()
tab_bottom <- plot_df |>
    arrange(desc(econ_mean)) |>
    head() |>
    arrange(econ_mean)
tab_middle <- plot_df |>
    arrange(econ_mean) |>
    slice(305:310)

tab <- bind_rows(tab_top, tab_middle, tab_bottom) |>
    mutate(DisplayName = sub(" and ", " & ", DisplayName)) |>
    mutate(DisplayName = sub(" \\(.*", "", DisplayName)) |>
    dplyr::select(MP = DisplayName, Party = Party.24,
                  `Mean` = econ_mean, `Lo` = `econ_lo`, `Hi` = `econ_hi`,
                  `Rank` = actual_rank, `Lo ` = `lo_rank`, `Hi ` = `hi_rank`)

kbl(tab,
    digits = 1,
    booktabs = TRUE) |>
    add_header_above(c(" " = 2, "Position" = 3, "Rank" = 3)) |>
    pack_rows(index = c("Left-wing" = 6, "Middle" = 6, "Right-wing" = 6))

```

The MPs with the furthest left positions are a mix of female MPs from
ethnic minorities and older white men. Nadia Whittome, Bell
Ribeirro-Addy and Diane Abbott all have some possibility of being *the*
most left-wing MP, and we can be confident that they are amongst the
top twenty most left-wing. Ian Byrne, John McDonnell and Jon Trickett
are slightly further to the right, and in their case we could only
confidently say that they were amongst the thirty most left-wing MPs. 

The rank order of these MPs seems *prima facie* plausible. All are
members of the Socialist Campaign Group. Ian Byrne, whilst elected as
a Labour MP, currently sits as an independent, having had the whip
withdrawn after voting to end child benefit caps.

We can be reasonably confident that the individuals at the top of
@tbl-extremes are to the left of the 2024- parliament, but we cannot
be so confident about the individuals shown in the middle of
@tbl-extremes. These are individuals who are close to the middle of
the distribution within the Parliament, but who, given the scale of
Labour's victory, need not be "centrist" as that term is commonly
understood. The degree of uncertainty for these individuals is
considerable. If we take the newly elected MP for Harlow Chris Vince
as an example -- we can rule out his being amongst the 100 most
left-wing MPs, but can't rule out his being amongst the most
right-wing Labour MPs. 

Some degree of certainty returns when we look at individuals furthest
to the right -- two MPs from Reform UK, and four conservatives. Four
of these MPs -- Farage, Lowe, Duncan Smith and Braverman -- have a
non-zero probability of being the most right-wing of MPs to feature in
this data; the other two are not far behind. The positioning of Reform
UK MPs is not surprising, and Braverman has been discussed as a
potential detection to Reform UK following her husband's defection
(which took place after the survey closed, and could therefore not
have affected respondents' evaluation of Braverman).


```{r}
#| fig-width: 9
#| fig-height: 6
#| label: fig-spcult
#| fig-cap: "Scatter plot of MPs' left-right positions on the economic dimension against position on the cultural dimension, for MPs elected in 2024. "

plot_df <- dat |>
    mutate(Party = ifelse(is.na(Party.24), Party.22, Party.24),
           Party = dplyr::recode(Party,
                                 "Conservative" = "Con",
                                 "Labour" = "Lab",
                                 "Plaid Cymru" = "PC",
                                 "Scottish National Party" = "SNP")) |>
    filter(Party != "Spk") |>
    filter(Party != "Ind") |>
    mutate(Party = factor(Party, ordered = TRUE),
           Party = reorder(Party, econ_mean))


ggplot(plot_df,
       aes(x = econ_mean, y = cult_mean)) +
    geom_point(aes(fill = Party, colour = Party, shape = Party),
               size = 3,
               alpha = 1/2) +
    scale_x_continuous("Estimated left-right position on the economic dimension") +
    scale_y_continuous("Estimated left-right position on the cultural dimension") +
    scale_shape_manual("Party",
                      values = c("Con" = 21,
                                 "Lab" = 22,
                                 "LD" = 23,
                                 "SNP" = 24,
                                 "PC" = 25,
                                 "Green" = 21,
                                 "RUK" = 22)) +
    scale_fill_manual("Party",
                  values = c("Con" = "#0087DC",
                             "Lab" = "#E4003B",
                             "LD" = "#FAA61A",
                             "SNP" = "#FDF38E",
                             "PC" = "#005B54",
                             "Green" = "#528D6B",
                             "RUK" = "#12B6CF")) +
    scale_colour_manual("Party",
                      values = c("Con" = "darkgrey",
                                 "Lab" = "darkgrey",
                                 "LD" = "black",
                                 "SNP" = "black",
                                 "PC" = "darkgrey",
                                 "Green" = "darkgrey",
                                 "RUK" = "black"
                                 )) +
    geom_smooth(method = "lm", se = FALSE) +
    theme_bw(base_size = 12) +
    theme(legend.position = "bottom")

dims_correl <- cor(plot_df$econ_mean, plot_df$cult_mean)
dims_correl <- round(dims_correl, 3)
```



```{r}
#| label: tbl-caucus
#| tbl-cap: Average positions and standard deviations of MPs in different party caucuses, 2019 to 2024 parliament. 

old <- read.csv(here::here("data",
                           "previous_exercise_with_wikidata_idents.csv"))
plot_df <- left_join(dat,
                     old,
                     by = join_by(wikidata))

cons_grp <- plot_df |>
    filter(Party.22 == "Conservative") |>
    summarize(nobs = n(), medianval = median(econ_mean), sdval = sd(econ_mean)) |>
    mutate(Group = "(All Conservative MPs)")

cons_onenation <- plot_df |>
    filter(Party.22 == "Conservative") |>
    filter(OneNation == 1) |>
    summarize(nobs = n(), medianval = median(econ_mean), sdval = sd(econ_mean)) |>
    mutate(Group = "One Nation members")


cons_trg <- plot_df |>
    filter(Party.22 == "Conservative") |>
    filter(TRG_Patron == 1) |>
    summarize(nobs = n(), medianval = median(econ_mean), sdval = sd(econ_mean)) |>
    mutate(Group = "Tory Reform Group")

cons_erg <- plot_df |>
    filter(Party.22 == "Conservative") |>
    filter(ERG_Subscriber == 1) |>
    summarize(nobs = n(), medianval = median(econ_mean), sdval = sd(econ_mean)) |>
    mutate(Group = "ERG subscribers")


cons_grp <- bind_rows(list(cons_grp, cons_onenation, cons_trg, cons_erg)) |>
    arrange(desc(medianval)) |>
    mutate(Party.22 = "Conservative") |>
    dplyr::select(Group, `N` = nobs, `Median` = medianval, `SD` = sdval)


###
lab_grp <- plot_df |>
    filter(Party.22 == "Labour") |>
    summarize(nobs = n(), medianval = median(econ_mean), sdval = sd(econ_mean)) |>
    mutate(Group = "(All Labour MPs)")

lab_scg <- plot_df |>
    filter(Party.22 == "Labour") |>
    filter(SocialistCampaignGroup == 1) |>
    summarize(nobs = n(), medianval = median(econ_mean), sdval = sd(econ_mean)) |>
    mutate(Group = "Socialist Campaign Group")


lab_trib <- plot_df |>
    filter(Party.22 == "Labour") |>
    filter(TribuneGroup == 1) |>
    summarize(nobs = n(), medianval = median(econ_mean), sdval = sd(econ_mean)) |>
    mutate(Group = "Tribune Group")

lab_grp <- bind_rows(list(lab_grp, lab_scg, lab_trib)) |>
    arrange(desc(medianval)) |>
    mutate(Party.22 = "Labour") |>
    dplyr::select(Group, `N` = nobs, `Median` = medianval, `SD` = sdval)


require(kableExtra)
tab <- rbind(cons_grp, lab_grp) |> arrange(desc(Median))
kbl(tab,
    booktabs = TRUE,
    digits = 2) |>
    pack_rows(index = c("Cons." = 4, "Lab." = 3))

```

*Third,* the positions of different factions within the two largest
parties (according to the MPs who are members of those factions) make
sense. We have already touched on this in our analysis of the
furthest-left individuals in the Labour Party, all of whom are members
of the Socialist Campaign Group. @tbl-caucus shows summary statistics
for two Labour factions and three Conservative factions, for members
of the 2019-2024 parliament.[^olddata] For Labour, Socialist Campaign
Group MPs are to the left of the party generally, and the party
generally is to the left of Tribune Group MPs. For the Conservatives,
members of the (partially overlapping) Tory Reform and One Nation
groupings are to the left of the party generally, and the party
generally is to the left of European Reform Group subscribers. This
suggests that the estimates are not just correct at the extremes shown
in @tbl-extremes, but capture differences between relevant groups of
MPs.

[^olddata]: We use data on the 2019 to 2024 parliament rather than the
    2024 - parliament because new members may still be establishing
    their membership of different party factions, and because there
    are now many fewer surviving members of the Tory Reform Group or
    One Nation. 

*Fourth,* MPs' positions on the economic dimension are almost identical
to their positions on the cultural dimension. @fig-bees and
@tbl-extremes both showed MPs' positions on the *economic* left-right
dimension. Had they instead showed positions on the cultural
left-right dimension, they would not have been very
different. @fig-spcult shows the relationship between MPs' positions
on the economic left-right dimension (horizontal axis) against their
positions on the social and cultural dimension (vertical axis). There
is a very strong relationship between positions on the two dimensions
($r=$ `r dims_correl`). This suggests that MPs' positions on these
two dimensions cannot be distinguished by knowledgeable observers like
councillors. To the extent that there are differences between
positions on these two dimensions, these differences are the result of
party-specific rather than individual-specific differences. Thus,
(most) Reform UK MPs are above and to the left of (most) Conservative
party MPs, indicating that Reform UK MPs are more right-wing on
cultural issues than they are on economic issues. Conversely, Liberal
Democrats are to the bottom and to the right of most Labour MPs,
suggesting that the Liberal Democrats are more right-wing on economic
issues than they are on cultural issues -- or equivalently, more
left-wing on cultural issues than they are on economic issues. The
relative positions of Reform UK and the Liberal Democrats make sense
given the stated ideologies of those parties; it is somewhat harder to
work out why SNP MPs should be judged more socially conservative than
most Labour MPs.


*Fifth,* MPs' positions in 2024 are almost identical to their
positions in 2023. When modelling responses we allowed MP positions to
be correlated across dimensions and connected across waves of the
survey. Having just shown that MPs' positions are near identical
across dimensions of ideology, it should not be surprising to learn
that MP positions are near identical when comparing estimates of
positions in wave 1 and estimates of positions in wave 2. The
correlation between the two positions is almost perfect: $r=1$ to
three decimal places. Although it is possible that some MPs may shift
in position over longer spans of time, there is no evidence of
shifting over this short period of one year.

```{r, eval = FALSE}
increments <- readRDS(file = here::here("working", "overtime_comparison.rds"))
increments <- increments |>
    group_by(wikidata, party) |>
    summarize(old_value.sc = mean(old_value.sc),
              value.sc = mean(value.sc))



ggplot(increments,
       aes(x = old_value.sc, y = value.sc)) +
    geom_point(aes(fill = party, colour = party, shape = party),
               size = 3,
               alpha = 1/2) +
    scale_x_continuous("Estimated left-right position on the economic dimension in 2023") +
    scale_y_continuous("Estimated left-right position on the cultural dimension in 2024") +
    scale_shape_manual("Party",
                      values = c("Con" = 21,
                                 "Lab" = 22,
                                 "LD" = 23,
                                 "SNP" = 24,
                                 "PC" = 25,
                                 "Green" = 21,
                                 "RUK" = 22)) +
    scale_fill_manual("Party",
                  values = c("Con" = "#0087DC",
                             "Lab" = "#E4003B",
                             "LD" = "#FAA61A",
                             "SNP" = "#FDF38E",
                             "PC" = "#005B54",
                             "Green" = "#528D6B",
                             "RUK" = "#12B6CF")) +
    scale_colour_manual("Party",
                      values = c("Con" = "darkgrey",
                                 "Lab" = "darkgrey",
                                 "LD" = "black",
                                 "SNP" = "black",
                                 "PC" = "darkgrey",
                                 "Green" = "darkgrey",
                                 "RUK" = "black"
                                 )) +
    geom_smooth(method = "lm", se = FALSE) +
    theme_bw(base_size = 12) +
    theme(legend.position = "bottom")

temporal_correl <- with(increments, cor(old_value.sc, value.sc))
temporal_correl <- round(temporal_correl, 3)
```





```{r}
### Read in Conor's estimates


twitter <- read.csv(here::here("data",
                               "mp_follower_ca_model_estimates.csv"))
dat <- left_join(dat,
                 twitter,
                 by = join_by(wikidata))

twitter_correl <- with(dat,
                       cor(econ_mean, svd.phi_1, use = "pairwise"))

twitter_N <- with(dat, length(which(!is.na(econ_mean) & !is.na(svd.phi_1))))
### Read in Conor's experts
experts <- read.csv(here::here("data",
                               "ideology_scores_expert_survey_aggregated.csv")) |>
    dplyr::select(wikidata, expert_mean = Mean)

dat <- left_join(dat,
                 experts,
                 by = join_by(wikidata))

expert_correl <- with(dat,
                      cor(econ_mean, expert_mean, use = "pairwise"))

expert_N <- with(dat, length(which(!is.na(econ_mean) & !is.na(expert_mean))))
### What, for sake of comparison, is the correlation between Conor's estimates and the experts?
foo <- left_join(twitter, experts, by = "wikidata")
cf_correl <- with(foo, cor(svd.phi_1, expert_mean, use = "pairwise"))

### Read in old EDM stuff
edm <- read.csv(here::here("data",
                           "hlv_lr.csv"))

lu <- read.csv(here::here("data",
                          "all_postwar_mps_in_twfy_all_wikidata_ids.csv")) |>
    dplyr::select(parlparseid = person_id,
                  wikidata = item) |>
    mutate(wikidata = sub("https", "http", wikidata),
           wikidata = sub("/wiki/", "/entity/", wikidata, fixed = TRUE))

edm <- left_join(edm,
                 lu,
                 by = join_by(parlparseid))

edm <- left_join(edm,
                 dat,
                 by = join_by(wikidata))

edm_correl <- with(edm,
                   cor(econ_mean, MP.position, use = "pairwise"))

edm_N <- with(edm, length(which(!is.na(econ_mean) & !is.na(MP.position))))

twitter_correl <- round(twitter_correl, 3)
expert_correl <- round(expert_correl, 3)
cf_correl <- round(cf_correl, 3)
edm_correl <- round(edm_correl, 3)
```

*Sixth,* the estimates are strongly associated with other estimates of
MPs' positions. @gaughan2024estimating estimates the positions of MPs
in the 2019 to 2024 parliament active on X (formerly Twitter) by
performing a latent correspondence analysis on their follower
matrix. The correlation between the estimates here and those estimates
derived from social media is high ($r=$ `r twitter_correl`, $N=$ 
`r twitter_N`). The correlation between the estimates here and the expert
survey used by Gaughan to validate his estimates is also high ($r=$ 
`r expert_correl`, $N=$ `r expert_N`) and indeed higher than the
correlation between Gaughan's social media estimates and the expert
survey ($r=$ `r cf_correl`, $N=$ `r expert_N`).

An earlier set of estimates comes from @hanretty2017dyadic who
(following @kellermann2012estimating) estimate the positions of
backbenchers in the 2010-2015 parliament by analysing the early day
motions they sign. The correlation with these estimates is also
extremely high ($r=$ `r edm_correl`, $N=$ `r edm_N`). 

There is therefore evidence from other attempts to estimate MPs'
positions to suggest that the estimates presented here display
convergent validity. We argue that our estimates are preferable to
these other estimates because of their greater coverage, and because
they display a stronger association with expert placements, which are
as close as we will get to "ground truth".

```{r uncertainty}
#| fig-width: 9
#| fig-height: 6
#| label: fig-uncertainty
#| fig-cap: "Uncertainty in estimates as a function of the number of comparisons"

plot_df <- dat |>
    rowwise() |>
    mutate(features_a = sum(grepl(wikidata, svy$wikidata.A, fixed = TRUE)),
           features_b = sum(grepl(wikidata, svy$wikidata.B, fixed = TRUE))) |>
    ungroup() |>
    mutate(n_comparisons = features_a + features_b,
           ci_width = econ_hi - econ_lo) |>
    mutate(type = case_when(is.na(PCON22CD) & !is.na(PCON24CD) ~ "New MP in 2024",
                            !is.na(PCON22CD) & is.na(PCON24CD) ~ "Left in 2024",
                            !is.na(PCON22CD) & !is.na(PCON24CD) ~ "Continuing MP"))




loglogmod <- lm(log(ci_width) ~ log(n_comparisons) + type, data = plot_df)

plot_df$residsq <- resid(loglogmod)^2
library(ggrepel)
ggplot(plot_df, aes(x = n_comparisons, y = ci_width)) +
    geom_point(aes(colour = type, shape = type)) +
    geom_text_repel(data = plot_df |> filter(residsq > .08),
              aes(label = DisplayName),
              hjust = 1,
              nudge_x = -0.025) + 
    scale_x_continuous("No. comparisons (log scale)",
                       trans = log_trans(),
                       breaks = c(1, 2, 5, 10, 20, 50, 250)) +
    scale_y_continuous("Width of 90% credible interval (log scale)",
                       trans = log_trans(),
                       breaks = c(2, 5, 10, 20)) +
    geom_smooth(aes(colour = type, linetype = type), se = FALSE) + 
    scale_shape_manual("Type", values = 0:2) +
    scale_colour_discrete("Type") +
    scale_linetype_discrete("Type", guide = "none") + 
    theme_bw(base_size = 12) +
    theme(legend.position = "bottom")
```


*Seventh,* uncertainty in the position of MPs is primarily a function of the number of times they were asked about and secondarily a function of whether they are new MPs. @fig-uncertainty shows the width of the 90% credible interval (logged, on the vertical axis) as a function of the number of comparisons (logged, on the horizontal scale). The three trend lines show nonlinear smooths for three groups of MPs: MPs who continued between both parliaments, MPs who left (voluntarily or otherwise) at the end of the 2019-2024 parliament, and MPs elected in the 2024 parliament but who did not sit in the 2019-2024 parliament. Although difficult to see on the graph, the trend line for "new" entrants is almost always higher than the trend line for the other two types of MPs.

# Conclusion

In this note we have described a method that we use for inferring the
left-right positions of MPs from a survey of local councillors. This
method, applied to this data, means that we can provide estimates of
positions for almost all MPs, estimates which are informed by local
knowledge and which come with estimates of uncertainty. These
estimates will be useful for researchers who are interested in MPs'
positions as an outcome (of, variously, initial selection, subsequent
re-election, and possible de-selection) and MPs' positions as
predictors of other outcomes (for example: cabinet recruitment and
declarations of support in leadership contests). 

We believe that these estimates of MP positions are superior to other
existing estimates of MP positions. Estimates of MP positions based on
co-sponsorship of early day motions are possible only for backbench
MPs, and suffer from the problem that many fewer MPs now sign early
day motions. Estimates from social media are only available for those
MPs who are active on a specific social media platform, and exhibit
lower correlation with expert placements of MPs. 

This method is not perfect. It cannot be applied retrospectively and
like any survey-based research it costs money. In addition when
applied to a single parliamentary term -- particularly a parliamentary
term where there are a large number of new MPs -- the degree of
uncertainty may be very large indeed, such that we cannot discriminate
between the positions of most new entrants. When applied to a single
parliamentary term, there are also issues with regional isolates --
areas where most MPs are uniformly to the left or right of the anchor
MPs, and where other MPs appear as very left- or right-wing as a
result. This issue arises in Scotland, where many MPs are to the left
of both Keir Starmer and Rishi Sunak, and where Liberal Democrat and
Conservative MPs appear very right-wing as a result, and in Essex,
where many MPs are to the right of both Keir Starmer and Rishi Sunak,
and where other MPs, including some Conservative MPs, can appear more
left-wing than they may in fact be.

We have motivated this note on the basis that these estimates can be
*useful*. Whilst contributing useful data is important (and worthy of
publication) we also think that our survey suggests some substantive
conclusions. Specifically, it suggests that MPs' positions on the
economic and sociocultural dimensions are so highly correlated that
knowledgeable observers position the average MP in the same place on
both dimensions. This does not imply that for any specific MP their
positions on those dimensions are identical. For any high profile MP
with a large public record, it is possible that knowledgeable
observers might be able to identify different positions on both those
domains. These individuals are, by definition, not average. More
generally, for any MP it is possible that the MP does *in fact* have
different positions on these dimensions, but that they have not said
or done enough which would lead a knowledgeable observer to identify
these contrasting positions correctly. As always, we can only infer
positions or ideal points from what MPs do or say; no social science
method is capable of peering into their soul. 

# References
